Publilia_2019_CHAPTER-1-strategies

### 2023-08-24 scripts/old/Publilia_2019_CHAPTER-1-strategies.Rmd
added old scripts to `scripts/old/` for reference

---
title: '01 - TreehopperFitness: Defining strategies and calculating frequencies'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 6)

library(kableExtra)   ## Needed if rendering tables
```

```{=html}
<style type="text/css">
  .book .book-body .page-inner {
    max-width: 1600px;
    margin-left: auto;
    margin-right: auto;
  }
details > summary {
  display: list-item;
}
</style>
```

# Publilia 2019 Chapter - Script 1: guarding strategies

## Required libraries and functions
```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggh4x)
library(rstatix)
library(emmeans)
library(effectsize)
library(gt)
library(gtsummary)
theme_gtsummary_compact()
library(flextable)
library(multcompView)
library(here)

@@

# 0 - Data import

clutch_status_guarding_summary_v2
```{r}
All_data_v6 <- read_rds(here("saved_rds/All_data_v6.rds"))
# clutch_status_guarding_summary_v2 <- read_rds(here("saved_rds/clutch_status_guarding_summary_v2.rds"))
# last_day_fID <- read_rds(here("saved_rds/last_day_fID.rds"))

strats_data <- read_rds(here("saved_rds/strats_data.rds"))
```

@@

## ****does this change anything?*****
paste("Including these missing females in the analysis reveals...")
```

# 2 - When were females visited on their first clutch?
First I need to know when females were visited on their first clutch.
I also need to estimate the hatch date for clutches that never hatched so I can still show and run stats for relative desertion date. I do this by adding 20 days to the initiation date. The mean interval from initiation to hatch is 20 days.
```{r}
## Moved to Publilia-01-qc.Rmd
# visited_1st_4d_data <- strats_data %>%
#   group_by(location) %>%
#     mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
#   ungroup() %>%
#   mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
#                                       "Visited", "Not visited"),
#          ants = ifelse(ants_removed, "Ants removed", "Control"),
#          est_hatch = as_date(ifelse(is.na(hatch_date),
#                                  initiation_date + 20,
#                                  hatch_date)),
#          est_rel2h = visit_ends-est_hatch)

@@

I also need to know how to pick just one initiator when two (or more) females are present on a clutch on the first day eggs are observed.
```{r}
## Moved to Publilia-01-qc.Rmd
# visited_1st_4d_data_tb <- visited_1st_4d_data %>%
#   # 1) Longest initiator
#   mutate(initL = init_first_dur_rank == 1) %>% 
#   # 2) Shortest initiator
#   #  i) break ties by picking marked females over unmarked
#   group_by(location, init_first_dur_shortest) %>% 
#     mutate(initS_tmp = ifelse(n()>1, !grepl("Uf", Female_ID), TRUE)) %>% 
#   ungroup() %>%
#   #  ii) finally picking at random (only one case, identical predictor vars)
#   group_by(location, init_first_dur_shortest, initS_tmp) %>%
#     mutate(initS = init_first_dur_shortest & initS_tmp & row_number()==1) %>%
#   ungroup()
```

# 3 - Defining exposure to ants

## 3.1 - Ant counts on plant; some ants_removed clutches were visited by ants, and vice versa
```{r}
## Moved to Publilia-01-qc.Rmd
# ant_counts <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch) %>%
#     summarize(ants_removed=first(ants_removed),
#               females_removed=first(females_removed),
#               n_antsP=sum(ants_P,na.rm=T),
#               ever_antsP=n_antsP>0,
#               mean_antsP=mean(ants_P,na.rm=T),
#               med_antsP=median(ants_P,na.rm=T),
#               
#               n_antsP_b4h=sum(ants_P&b4h),
#               ever_antsP_b4h=n_antsP_b4h>0,
#               mean_antsP_b4h=mean(ifelse(b4h,ants_P,NA),na.rm=T),
#               med_antsP_b4h=median(ifelse(b4h,ants_P,NA),na.rm=T),
#               
#               n_antsP_afh=sum(ants_P&afh),
#               ever_antsP_afh=n_antsP_afh>0,
#               mean_antsP_afh=mean(ifelse(afh,ants_P,NA),na.rm=T),
#               med_antsP_afh=median(ifelse(afh,ants_P,NA),na.rm=T)) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

@@

## 3.2 - Ant arrival times
```{r}
## Moved to Publilia-01-qc.Rmd
# antP_arrival_afh_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsP_afh=ants_P&afh) %>%
#     filter(is_antsP_afh) %>%
#     slice_head(n=1) %>% 
#   mutate(antP_arrival_afh=Date-est_hatch) %>%
#   select(location, antP_arrival_afh, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# antP_last_b4h_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsP_bfh=ants_P&b4h) %>%
#     filter(is_antsP_bfh) %>%
#     slice_tail(n=1) %>% 
#   mutate(antP_last_b4h=Date-est_hatch) %>%
#   select(location, antP_last_b4h, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# antL_arrival_afh_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsL_afh=ants_L&afh) %>%
#     filter(is_antsL_afh) %>%
#     slice_head(n=1) %>% 
#   mutate(antL_arrival_afh=Date-est_hatch) %>%
#   select(location, antL_arrival_afh, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# antL_last_b4h_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsL_bfh=ants_L&b4h) %>%
#     filter(is_antsL_bfh) %>%
#     slice_tail(n=1) %>% 
#   mutate(antL_last_b4h=Date-est_hatch) %>%
#   select(location, antL_last_b4h, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# ant_data <- ant_counts %>%
#   left_join(antP_last_b4h_data) %>%
#   left_join(antP_arrival_afh_data) %>%
#   left_join(antL_last_b4h_data) %>%
#   left_join(antL_arrival_afh_data)
# 
# visited_1st_4d_data_ants <- visited_1st_4d_data %>%
#   left_join(ant_data)

strats_data %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h, y=antL_last_b4h, color=females_removed)) +
  geom_point() +
  geom_abline(slope=1)

strats_data %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h, y=antL_arrival_afh, color=females_removed)) +
  geom_point() +
  geom_smooth(method=lm) +
  geom_abline(slope=1)

strats_data %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_last_b4h)) +
  geom_boxplot()

strats_data %>%
  filter(ants=="Control") %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_arrival_afh)) +
  geom_boxplot() +
  geom_jitter(pch = 21, size=1, width=0.35, height=0) +
  stat_summary(fun=mean, geom="point", shape=4, size=3)

strats_data %>%
  filter(ants=="Control", !females_removed) %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_arrival_afh)) +
  geom_boxplot() +
  geom_jitter(pch = 21, size=1, width=0.35, height=0) +
  stat_summary(fun=mean, geom="point", shape=4, size=3)
```

# 4 - Figure 1: Histograms of guarding tenure

## 4.1 Using observed presence of ants as ants treamment

Two proxies:

1. ants ever observed on plant before hatch `hatchever_antsP_b4h` or 
2. ants ever observed on plant after hatch `ever_antsP_afh`

### Not visited early
```{r}
## Not visited early
# ever ants b4h
ggplot(data = filter(strats_data, 
                                 `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends<last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color = ever_antsP_b4h, fill = ever_antsP_b4h), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)

# ever ants afh
ggplot(data = filter(strats_data, 
                                 `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color=ever_antsP_afh, fill=ever_antsP_afh), 
            alpha = 0.75, stat='bin', binwidth=4, size=0, position="identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)
```

### Visited early
```{r}
# ever ants b4h
ggplot(data = filter(strats_data,
                     `visited in 4 days` == "Visited",
                     !grepl("U", Female_ID), !females_removed, mom_at_init,
                     clutch_initiated_order == 1, visit_ends<last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color = ever_antsP_b4h, fill = ever_antsP_b4h), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)

# ever ants afh
ggplot(data = filter(strats_data,
                     `visited in 4 days` == "Visited",
                     !grepl("U", Female_ID), !females_removed, mom_at_init,
                     clutch_initiated_order == 1, visit_ends < last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color=ever_antsP_afh, fill=ever_antsP_afh), 
            alpha = 0.75, stat='bin', binwidth=4, size=0, position="identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)
```

## 4.2 - Using ant treatment (not observed presence)

#### Desertion relative to hatch
Using geom_area and density
```{r}
Fig1_v1_data <- visited_1st_4d_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID)

@@

# strat_data_table1 %>%
#   as_flex_table() %>%
#   flextable::save_as_docx(path=here("prelim_figs/strat_data_table1.docx"))

@@

# strat_data_table2 %>%
#   as_flex_table() %>%
#   flextable::save_as_docx(path=here("prelim_figs/strat_data_table2.docx"))

@@

#### Guarding duration instead of relative guarding duration
```{r}
Fig1_v1A.gd <- ggplot(data = filter(Fig1_v1_data, `visited in 4 days` == "Not visited")) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof initiating females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

@@

## 4.3 - Dealing with non-independence of multi-initiators: Using only longest or shortest initiator
Sample sizes
```{r}
# Samples sizes for "not visited" are the same across longests and shortests
filter(strats_data, `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initL) %>% nrow()

filter(strats_data, `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initS) %>% nrow()

# Shortests have more females than longests for some reason??
filter(strats_data, `visited in 4 days` == "Visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initL) %>% nrow()

filter(strats_data, `visited in 4 days` == "Visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initS) %>% nrow()

strats_data %>% filter(initL, mom_at_init, num_at_init>1) %>% nrow() # 34
strats_data %>% filter(initS, mom_at_init, num_at_init>1) %>% nrow() # 36

strats_data %>% filter(initL, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed) %>% nrow() # 13

strats_data %>% filter(initS, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed) %>% nrow() # 12

strats_data %>% filter(initL, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed,
                                  clutch_initiated_order == 1) %>% nrow() # 12

strats_data %>% filter(initS, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed,
                                  clutch_initiated_order == 1) %>% nrow() # 11

strats_data %>% filter(initL, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed,
                                  clutch_initiated_order == 1, visit_ends < last_day_fID) %>% nrow() # 5

>>>>

---
title: "Publilia_2019_CHAPTER-1-strategies"
author: "micah fletcher"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Publilia 2019 Chapter - Script 1: guarding strategies
This is meant to consolidate previous scripts and provide a readable and reproducible record of the analyses for the Publilia Chapter.

(See `Publilia_2019_CHAPTER-0-prep.Rmd` for the full outline of the chapter)
1) *Guarding strategies:* Three distinct strategies for maternal care emerge from ant-dependent and secondary visitor-dependent plasticity
  - *Dataset A* requirements:
    - Guarding durations (relative to [estimated] hatching) are accurate.
      - The sample of clutches is representative of the full factorial of ants / no ants and visited early / not visited early
    - Able to compare 
      - initiated vs. visited
      - first vs. later [initiated] clutches
        -however, if plotting "all clutches" I can include unmarked females
      - disappeared vs. deserted (based on last day a female was observed)
    - Deciding on who is the "true" initiator is necessary for assumptions of independence for statistics. 
      - Ultimately, comparing the two most obvious options (first and longest guarding mom at initiation, vs first and shortest guarding mom at initiation) will need to suffice.
Missing data is clearly represented; Compare the "original" unfiltered data to the final data and report why (and how many) data were excluded.

## Required libraries and functions
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(emmeans)
library(effectsize)

library(gt)
library(gtsummary)
library(flextable)
theme_gtsummary_compact()

library(multcompView)

@@

# 0 - Data import
clutch_status_guarding_summary_v2
```{r}
setwd("/Users/micah/Documents/Princeton/Kocher_Lab/Publilia/PubliliaBehavEcol/Watershed_2019/Watershed_2019_analysis/CHAPTER")

# clutch_status_guarding_summary_v2 <- read_csv("../Publilia_2019_ABS_talk_analysis_data/clutch_status_guarding_summary_v2")
# last_day_fID <- read_csv("../Publilia_2019_ABS_talk_analysis_data/last_day_fID")

strats_data <- left_join(clutch_status_guarding_summary_v2, last_day_fID)
```

@@

## ****does this change anything?*****
paste("Including these missing females in the analysis reveals...")

# 2 - Histograms and related stats
First I need to know when females were visited on their first clutch.
I also need to estimate the hatch date for clutches that never hatched so I can still show and run stats for relative desertion date. I do this by adding 20 days to the initiation date. The mean interval from initiation to hatch is 20 days.
```{r}
visited_1st_4d_data <- strats_data %>%
  group_by(location) %>%
    mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
  ungroup() %>%
  mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
                                      "Visited", "Not visited"),
         ants = ifelse(ants_removed, "Ants removed", "Control"),
         est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         est_rel2h = visit_ends-est_hatch)

@@

I also need to know how to pick just one initiator when two (or more) females are present on a clutch on the first day eggs are observed.
```{r}
visited_1st_4d_data_tb <- visited_1st_4d_data %>%
  # 1) Longest initiator
  mutate(initL = init_first_dur_rank == 1) %>% 
  # 2) Shortest initiator
  #  i) break ties by picking marked females over unmarked
  group_by(location, init_first_dur_shortest) %>% 
    mutate(initS_tmp = ifelse(n()>1, !grepl("Uf", Female_ID), TRUE)) %>% 
  ungroup() %>%
  #  ii) finally picking at random (only one case, identical predictor vars)
  group_by(location, init_first_dur_shortest, initS_tmp) %>%
    mutate(initS = init_first_dur_shortest & initS_tmp & row_number()==1) %>%
  ungroup()
```

# Ant counts on plant; some ants_removed clutches were visited by ants, and vice versa
```{r}
ant_counts <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch) %>%
    summarize(ants_removed=first(ants_removed),
              females_removed=first(females_removed),
              n_antsP=sum(ants_P,na.rm=T),
              ever_antsP=n_antsP>0,
              mean_antsP=mean(ants_P,na.rm=T),
              med_antsP=median(ants_P,na.rm=T),

              n_antsP_b4h=sum(ants_P&b4h),
              ever_antsP_b4h=n_antsP_b4h>0,
              mean_antsP_b4h=mean(ifelse(b4h,ants_P,NA),na.rm=T),
              med_antsP_b4h=median(ifelse(b4h,ants_P,NA),na.rm=T),

              n_antsP_afh=sum(ants_P&afh),
              ever_antsP_afh=n_antsP_afh>0,
              mean_antsP_afh=mean(ifelse(afh,ants_P,NA),na.rm=T),
              med_antsP_afh=median(ifelse(afh,ants_P,NA),na.rm=T)) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

@@

ant arrival times
```{r}
antP_arrival_afh_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsP_afh=ants_P&afh) %>%
    filter(is_antsP_afh) %>%
    slice_head(n=1) %>% 
  mutate(antP_arrival_afh=Date-est_hatch) %>%
  select(location, antP_arrival_afh, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

antP_last_b4h_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsP_bfh=ants_P&b4h) %>%
    filter(is_antsP_bfh) %>%
    slice_tail(n=1) %>% 
  mutate(antP_last_b4h=Date-est_hatch) %>%
  select(location, antP_last_b4h, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

antL_arrival_afh_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsL_afh=ants_L&afh) %>%
    filter(is_antsL_afh) %>%
    slice_head(n=1) %>% 
  mutate(antL_arrival_afh=Date-est_hatch) %>%
  select(location, antL_arrival_afh, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

antL_last_b4h_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsL_bfh=ants_L&b4h) %>%
    filter(is_antsL_bfh) %>%
    slice_tail(n=1) %>% 
  mutate(antL_last_b4h=Date-est_hatch) %>%
  select(location, antL_last_b4h, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

ant_data <- ant_counts %>%
  left_join(antP_last_b4h_data) %>%
  left_join(antP_arrival_afh_data) %>%
  left_join(antL_last_b4h_data) %>%
  left_join(antL_arrival_afh_data)

visited_1st_4d_data_ants <- visited_1st_4d_data %>%
  left_join(ant_data)

visited_1st_4d_data_ants %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h, y=antL_last_b4h, color=females_removed)) +
  geom_point() +
  geom_abline(slope=1)

visited_1st_4d_data_ants %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h, y=antL_arrival_afh, color=females_removed)) +
  geom_point() +
  geom_smooth(method=lm) +
  geom_abline(slope=1)

visited_1st_4d_data_ants %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_last_b4h)) +
  geom_boxplot()

visited_1st_4d_data_ants %>%
  filter(ants=="Control") %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_arrival_afh)) +
  geom_boxplot() +
  geom_jitter(pch = 21, size=1, width=0.35, height=0) +
  stat_summary(fun=mean, geom="point", shape=4, size=3)

visited_1st_4d_data_ants %>%
  filter(ants=="Control", !females_removed) %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_arrival_afh)) +
  geom_boxplot() +
  geom_jitter(pch = 21, size=1, width=0.35, height=0) +
  stat_summary(fun=mean, geom="point", shape=4, size=3)

```{r}
## Not visited early
# ever ants b4h
ggplot(data = filter(visited_1st_4d_data_ants, 
                                 `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends<last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color = ever_antsP_b4h, fill = ever_antsP_b4h), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)

# ever ants afh
ggplot(data = filter(visited_1st_4d_data_ants, 
                                 `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color=ever_antsP_afh, fill=ever_antsP_afh), 
            alpha = 0.75, stat='bin', binwidth=4, size=0, position="identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)

## Visited early
# ever ants b4h
ggplot(data = filter(visited_1st_4d_data_ants,
                     `visited in 4 days` == "Visited",
                     !grepl("U", Female_ID), !females_removed, mom_at_init,
                     clutch_initiated_order == 1, visit_ends<last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color = ever_antsP_b4h, fill = ever_antsP_b4h), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)

# ever ants afh
ggplot(data = filter(visited_1st_4d_data_ants,
                     `visited in 4 days` == "Visited",
                     !grepl("U", Female_ID), !females_removed, mom_at_init,
                     clutch_initiated_order == 1, visit_ends < last_day_fID)) +
  geom_area(aes(x = est_rel2h, y=..density.., 
                color=ever_antsP_afh, fill=ever_antsP_afh), 
            alpha = 0.75, stat='bin', binwidth=4, size=0, position="identity") + 
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Proportion of females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1)
```

## Figure 1:
Using geom_area and density
```{r}
Fig1_v1_data <- visited_1st_4d_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID)

@@

strat_data_table1 %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="strat_data_table1.docx")

@@

strat_data_table2 %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="strat_data_table2.docx")

@@

Guarding duration instead of relative guarding duration
```{r}
Fig1_v1A.gd <- ggplot(data = filter(Fig1_v1_data, `visited in 4 days` == "Not visited")) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof initiating females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

@@

*Dealing with non-independence of multi-initiators: Using only longest or shortest initiator*
Sample sizes
```{r}
# Samples sizes for "not visited" are the same across longests and shortests
filter(visited_1st_4d_data_tb, `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initL) %>% nrow()

filter(visited_1st_4d_data_tb, `visited in 4 days` == "Not visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initS) %>% nrow()

# Shortests have more females than longests for some reason??
filter(visited_1st_4d_data_tb, `visited in 4 days` == "Visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initL) %>% nrow()

filter(visited_1st_4d_data_tb, `visited in 4 days` == "Visited",
                                 !grepl("U", Female_ID), !females_removed, mom_at_init,
                                 clutch_initiated_order == 1, visit_ends < last_day_fID,
                                 initS) %>% nrow()

visited_1st_4d_data_tb %>% filter(initL, mom_at_init, num_at_init>1) %>% nrow() # 34
visited_1st_4d_data_tb %>% filter(initS, mom_at_init, num_at_init>1) %>% nrow() # 36

visited_1st_4d_data_tb %>% filter(initL, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed) %>% nrow() # 13

visited_1st_4d_data_tb %>% filter(initS, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed) %>% nrow() # 12

visited_1st_4d_data_tb %>% filter(initL, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed,
                                  clutch_initiated_order == 1) %>% nrow() # 12

visited_1st_4d_data_tb %>% filter(initS, mom_at_init, num_at_init>1,
                                  !grepl("U", Female_ID), !females_removed,
                                  clutch_initiated_order == 1) %>% nrow() # 11
